---
title: "Data Visualization"
author: "Wenxin Tian"
date: "`r Sys.Date()`"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)
```

## Load Weather Data

Using the rnoaa package to directly load data from the web

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

weather_df
```

## Scatterplots

Create a first scatter plot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

New approach using pipes:

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

Save and edit plot objects:

```{r}
weather_plot =
  weather_df |>
  ggplot(aes(x = tmin, y = tmax))

weather_plot + geom_point()
```

## Advanced Scatterplot

Start with the same one and make it fancy!

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point() +
  geom_smooth(se = FALSE)
```

What about `aes` placement?

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name)) + # Putting color here means color is only applied to the scatterplot and will not be used when drawing smooth line later on.
  geom_smooth(se = FALSE)
```

Facet:

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name) # . means nothing defies rows, and ~ means what defiens cols.

weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_grid(name ~ .)

# Define transparency (alpha) to make line more visible:

weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.2) +
  geom_smooth(se = FALSE, size = 2) +
  facet_grid(. ~ name)
```

Lets combine some elements and try a new plot:

```{r}
weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp, alpha = .5)) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

## Some small notes

How many geoms have to exist?
You can have whatever geoms you want.

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_smooth(se = FALSE) # only line, no points 
  
```

You can also use a neat geom:

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex() +
  geom_density2d()
```

## Univariate Plots

Histograms:

```{r}
weather_df |>
  ggplot(aes(x = tmin, color = name)) +
  geom_histogram()
```

Position not great, change position and change color to fill:

```{r}
weather_df |>
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram(position = 'dodge')
```

Put them side by side:

```{r}
weather_df |>
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram() +
  facet_grid(. ~ name)
```

Try a new geometry!

```{r}
weather_df |>
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = 0.3)
```

What about box plots?

```{r}
weather_df |>
  ggplot(aes(x = name, y = tmin)) + 
  geom_boxplot()
```

Other trendy plots (with statistics):

```{r}
weather_df |>
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin(alpha = 0.5) +
  stat_summary()
```

Ridge plots:

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = name, fill = name)) +
  geom_density_ridges(alpha = 0.5)
```

## Save and Embed

Save a scatterplot:

```{r}
weather_plot = 
  weather_df |>
    ggplot(aes(x = tmin, y = tmax, color = name)) +
    geom_point(alpha = 0.5)

ggsave("weather_plot.pdf", weather_plot, width = 5, height = 5)
```

Embedding same plot at different sizes:

```{r, fig.width=12, fig.height=6}
weather_plot
```


```{r, fig.width=6, fig.height=3}
weather_plot
```

